\documentclass[11pt]{article}

\setlength{\topmargin}{-.5in}
\setlength{\oddsidemargin}{0.0in}
\addtolength{\textwidth}{1.5in}
\addtolength{\textheight}{1.10in}
%\setlength{\textwidth}{6.0in}
%\setlength{\textheight}{8.0in}

\begin{document}

\noindent
{\Large\bf Gopalan Nadathur: Abbreviated Vita and Publications}\\

\medskip

\noindent {\large \bf Highest Degree}

\smallskip
\noindent Ph.D., Computer Science, University of Pennsylvania, May
1987

\noindent Thesis Title: A Higher-Order Logic as the Basis for Logic
Programming

\noindent Advisor: Dale Miller

\bigskip

\noindent {\large \bf Other Recent Appointments}

\medskip
\noindent None in the last five years


\bigskip\medskip
\noindent {\large \bf Selected Recent Publications}

\medskip
\noindent {\it [The expertise I bring to the project is that of using
    logical techniques to formalize and reason about program
    properties. The publications listed below are those from the last
    five years that indicate my familiarity with this area.]}

\bigskip
\noindent A modular approach to metatheoretic reasoning for extensible
  languages. {\it ACM Transactions on Programming Languages and
    Systems}, 47(3):1-56, September 2025. (With D. Michaelson and E. Van 
  Wyk.)

\medskip
\noindent Transporting theorems about typeability in LF across 
  schematically defined contexts. Proceedings of the {\it 24th
    International Symposium on Principles and Practice of Declarative
    Programming}, DOI 10.1145/3756907.3756919. Dec. 2025. (With
  C. Johnson.)

\medskip
\noindent Ground stratification for a logic of definitions with induction,
  {\it Proceedings of the Twentieth Workshop on Logical Frameworks and
    Meta-Languages: Theory and Practice}, Electronic Proceedings in
  Theoretical Computer Science (EPTCS), Vol 431: 1--16, Oct
  2025. (With N. Guermond.)

\medskip
\noindent Binding contexts as partitionable multisets in Abella.
   {\it Proceedings of the Nineteenth Workshop on Logical Frameworks
     and Meta-Languages: Theory and Practice}, Electronic Proceedings in
  Theoretical Computer Science (EPTCS), Vol 404, July 2024. (With T. Gray.)

\medskip
\noindent A logic for formalizing properties of LF
  specifications. Proceedings of the {\it 24th International Symposium
    on Principles and Practice of Declarative Programming}, DOI
  10.1145/3551357.3551377, Sept 2022. (With M. Southern.)

\medskip
\noindent Adelfa: A system for reasoning about LF specifications. {\it
  Proceedings of the Sixteenth Workshop on Logical Frameworks and
  Meta-Languages: Theory and Practice}, Electronic Proceedings in
  Theoretical Computer Science (EPTCS), Vol 337, July 2021. (With M. Southern.)

\medskip
\noindent A logic for reasoning about LF specifications. Manuscript
  available from the Computing Research Repository (CoRR), a
  subdivision of arXiv, at the URL
  \verb+https://arxiv.org/abs/2107.00111+, pages 1-72, June
  2021. (With M.~Southern.)
\end{document}
