\section*{Need Justification}


As explained in the cover letter and also manifest in the GIA category
chosen for this application, funds are being requested to support a
graduate research assistant whose work will help address some
questions raised with respect to a previous unsuccessful, but
competitive, National Science Foundation proposal.
%
% The particular area of work is not covered by the existing sponsored
% grants that are available to them. 

Our current cumulative non-sponsored funds total just over \$50,000.
%
Although these funds have not been officially committed to a specific
use, they are needed by the individual PIs to support aspects such as
travel (especially for students), equipment purchases, and visits by
collaborators related to their established research projects. Some of
these funds may be diverted to auxiliary aspects of this project, such
as to purchase computing equipment to be used in carrying out the
implementation and verification work. However, funds to support the
bulk of the requirement---the participation of a graduate research
assistant over an academic year---are not available. While the
investigators will look actively for other resources to foster work,
the GIA category they are applying to is in fact the first one that
they have found that is ideally designed for activities with the
orientation underlying this project.

PI Nadathur has access to one non-sponsored account, an Indirect Cost
Recovery (ICR) fund, of around \$15,000.
%
He is currently advising three doctoral students who are actively
writing and submitting papers to conferences. At this time, one of
these students is awaiting a response to a submitted paper and the
other two students intend to submit papers to conferences that will
occur in Fall 2026.
%
It is envisaged that around \$9000 will be needed
to cover their travel if all the papers are accepted.
%
Additionally, the Computer Science and Engineering department has
recently instituted a policy that requires faculty members to use ICR
funds to reduce their teaching load to one that is compatible with an 
active research program.
%
A commitment of \$2500 has been made towards this end for the 2026-27
academic year.
%
Accounting for travel and computing expenses for the student to be
supported by this grant (to be shared with Co-PI Van Wyk), will leave
this non-sponsored account with virtually no money to be used for this
project.

Co-PI Van Wyk has access to three non-sponsored accounts. These funds are
committed, to various degrees, to other projects or expenditures.
\begin{itemize}
\item Stipend for chairing Senate Committee on Academic Freedom and
   Tenure for 4 years: \$29,170.
   
   A large portion of these funds are committed to supporting the
   travel and computer needs of other graduate students that are
   supported on teaching assistantships. These three students will
   travel to conferences to present their published research work (and
   have already). Some of the primary venues for their work are
   primarily located in Europe, which adds to the cost.
   
   Over the next 2 years Van Wyk expects to spend at least \$10,000 on
   student travel.
   
   The students the PIs anticipate hiring with the funds proposed here will
   also incur travel costs to present their research results and will
   also need a computer on which to do their work. This will amount to
   around \$2,500.
   
   PI Van Wyk is chair of the steering committee for the ACM SIGPLAN
   International Conference on Software Language Engineering and uses
   the funds to travel to steering committee meetings. He also uses
   these funds for other research travel and this amounts to another
   \$5,000 over the next few years.
   
\item Indirect Cost Recovery (ICR) funds: \$3,238.

   As mentioned above, the Computer Science and Engineering department
   allows faculty members to buy out of the increase using ICR
   funds. Thus these funds are committed to buying out of courses to
   get to a teaching load compatible with an active research program.
   
\item IBM Gift: \$2,870

   These funds are part of a gift for investigating programmer tools
   for improving the ergonomics of using extensible languages and are
   thus committed to completing that project.
   
\end{itemize}
   

It is worth noting that the remaining uncommitted funds available to each PI are
rather limited and that in these challenging times for acquiring
research funds it is wise to keep hold on some funds for unanticipated
travel and students needs.
   
