\documentclass[11pt]{article}
\usepackage[letterpaper,margin=1.0in]{geometry}

\usepackage{url}
\usepackage[colorlinks=true,allcolors=black]{hyperref}

%\pagestyle{empty}

\begin{document}
%\centerline{\large Biographical Sketch}

\bigskip
\centerline{\Large\textbf{Eric Van Wyk}}

\bigskip
\noindent{\large\bf Education}
\medskip

\begin{tabular}{lll}
University of Iowa, Computer Science, Ph.D., 1998  \\
Dissertation: \emph{Semantic Processing for Macro Processors}  Adviser: Teodor Rus

\end{tabular}

\bigskip
\noindent\textbf{Appointments}
\medskip


\begin{tabular}{llll}
2021--pres.   & Professor & 
  Computer Science \& Eng. & University of Minnesota  \\
2008--2021   & Associate Professor & 
  Computer Science \& Eng. & University of Minnesota 
\end{tabular}

\bigskip
\noindent{\large\bf Recent Publications}

\begin{itemize}

\item Dawn Michaelson, Gopalan Nadathur, and Eric Van Wyk.
Modular Metatheory for Extensible Languages.
\emph{ACM Transactions of Programming Languages and Systems
(TOPLAS)}. 47(3). 1--56. September, 2025. 
% Submitted April, 2024, accepted March 2025.

% \url{doi.org/10.1145/3731555}



\item Luke Bessant, Eric Van Wyk.
Scheduling the Construction and Interrogation of Scope Graphs Using Attribute Grammars.
In \emph{Proc. of ACM SIGPLAN International Conference on Software
Language Engineering (SLE)}, pages 56--69, ACM Press, October 2025.
%\url{doi.org/10.1145/3732771.3742711}

\item Lucas Kramer, Eric Van Wyk.
Sharing Trees and Contextual Information: Re-imagining Forwarding in Attribute Grammars.
In \emph{Proc. of ACM SIGPLAN International Conference on Software Language Engineering (SLE)}.
pages 56--69,
ACM Press,
October, 2023. 
%\url{doi.org/10.1145/3623476.3623520}

\item Nathan Ringo, Lucas Kramer, and Eric Van Wyk.
Nanopass Attribute Grammars
In \emph{Proc. of ACM SIGPLAN International Conference on Software Language Engineering (SLE)}.
pages 70--83,
ACM Press,
October, 2023. 
%\url{doi.org/10.1145/3623476.3623514}


\item Lucas Kramer, Ted Kaminski, and Eric Van Wyk.
Reflection in Attribute Grammars: Implementation and Application.
\emph{Journal of Computer Languages.} 
101033. June, 2021.
Elsevier.


\end{itemize}


\bigskip
\noindent{\large\bf Older Relevant Publications}

\begin{itemize}


\item Dawn Michaelson and Eric Van~Wyk. Monadification of Attribute
  Grammars. In \emph{Proc. of ACM SIGPLAN Intl. Conf. on Software
    Language Engineering}, 2020, ACM. To appear.

%  \url{https://doi.org/10.1145/3426425.3426941}

\item Ted Kaminski, Lucas Kramer, Travis Carlson, and Eric Van
  Wyk. 2017. Reliable and Automatic Composition of Language Extensions
  to C: The \textsc{ableC} Extensible Language Framework. Proc.
  of the ACM Programming Languages, 1, OOPSLA, Article 98 (October,
  2017), 29 pages.

%  \url{https://doi.org/10.1145/3138224}

\item Ted Kaminski and Eric Van~Wyk.  Modular Well-Definedness
  Analysis for Attribute Grammars.  In \emph{Proc. of International
    Conference on Software Language Engineering}, volume 7745 of LNCS,
  pages 352--371, Springer-Verlag,
  2012.

%  \url{https://doi.org/10.1007/978-3-642-36089-3_20}.

\item August Schwerdfeger and Eric Van Wyk.  Verifiable Composition of
  Deterministic Grammars.  In \emph{Proc. of ACM SIGPLAN Conference on
    Programming Language Design and Implementation (PLDI)}.  ACM
  Press. June, 2009.

%  \url{http://doi.acm.org/10.1145/1542476.1542499}.

  

\end{itemize}


\end{document}
